\textbf{Designing a Virtual Security Layer for Cloud Content}

### 2015-12-01 tex/paper.tex
Restructured sections

% useful (wrappers for) math symbols:
\newcommand{\Cardinality}[1]{\left\lvert#1\right\rvert}
\newcommand{\Ceiling}[1]{\left\lceil#1\right\rceil}
\newcommand{\Floor}[1]{\left\lfloor#1\right\rfloor}
\newcommand{\Iff}{\Leftrightarrow}
\newcommand{\Implies}{\Rightarrow}
\newcommand{\Intersect}{\cap}
\newcommand{\Sequence}[1]{\left[#1\right]}
\newcommand{\Set}[1]{\left\{#1\right\}}
\newcommand{\SetComp}[2]{\Set{#1\SuchThat#2}}
\newcommand{\SuchThat}{\mid}
\newcommand{\Tuple}[1]{\langle#1\rangle}
\newcommand{\Union}{\cup}
\usetikzlibrary{positioning,shapes,shadows,arrows}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords: }} #1}

\title{\textbf{Designing a Virtual Security Layer for Cloud Content}}
\author{Lukas Klingsbo}

@@

\pagenumbering{roman}
\setcounter{page}{2}

@@

%}}}
\begin{abstract}
    TODO: Abstract

@@

%}}}
\section{Introduction}
Developing larger projects containing static content usually involves using a Content Distribution
Network to be able to scale to a large user base. The commercial Content Distribution Networks are
usually fairly easy to use, the content that is to be used in a project is usually simply uploaded
and then directly available to the public.  For secret content this can be a problem and that is
what this thesis is about. This work examines ways of enforcing access control on content and groups
of content in the form of folders and snapshots. A system was developed to make the underlying
theory work in practice.

The research question that this report answers is whether it is practically feasible to use
Copy-on-Write for a high-level system like the one that is implemented. The work also thoroughly
compares different types of lazy snapshots to figure out which is the most

\subsection{Prior Work}
\subsubsection{Copy-on-Write}
This work relies heavily on the Copy-on-Write principle which was founded and used in the Mach
kernel~\cite{COPYONWRITE} as it can be used to efficiently create snapshots.

Copy-on-Write is used in virtual memory management systems~\cite{VIRTCOW}, for snapshot
functionality in both file systems~\cite{BTRFS} and logical volume management systems~\cite{LVMCOW},
and as an optimisation technique for objects and types in programming languages\cite{LANGCOW}.

Its principle is that when processes or nodes share data in between each other, the data is not
copied until one of the processes does changes to it. This is an optimisation as the processes does
not have to send or copy all of the related data that is in memory, rather they only have to send
pointers to the data. After many Copy-on-Write's a complex structure can be built up, but it is
possible to solve that structure~\cite{COPYONWRITE2}.

TODO: Polish and extend

\newpage
\section{Related Terminology}
\subsection{Technologies}
\subsubsection{React}
React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and
the browser's, this makes it able to efficiently update dynamic web pages after a change of state
through comparing the old virtual DOM with the resulting virtual DOM after the state change and then
only update the browser's DOM according to the delta between the virtual DOMs~\cite{REACT}. React
can be seen as the system for handling views in front-ends implementing a MVC
(Model-View-Controller) architecture.

@@

In this work we use a RESTful API, which means that it is an API following the REST architecture.

@@

\subsection{Abbreviations}

\subsubsection{JPF}
Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
licence, which made more people contribute to the project. JPF is usually used for doing model
checking of concurrent programs to easily find for example race conditions and dead locks.

\subsection{Terms}
\subsubsection{Snapshot}
A snapshot is a way to record the full state of a system at a specific time. The term comes from
photography where a photo can be seen as the state of what the photo is of at a certain time.
Snapshots should not be confused with full copies of a system as full copies can be used as backups
meanwhile snapshots are not very effective means of backups in the case of data corruption. It is
not effective against data corruption as snapshots usually still refer to unchanged data that is
still a part of the system~\cite{SNAPSHOT}.

\section{Background}
\subsection{About Uprise}
Uprise is a company based in Uppsala, Sweden. It is an EA studio focussing on creating great gaming
experiences, that means that they are not focussed on the actual gameplay which other EA studios
like DICE is.

\subsection{The current system}
Today a system called battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files,
mostly images, to content delivery networks. The current system does not make use out of the
security features that the CDN's are offering, instead it uses a form of security by obscurity. When
a file is uploaded to a CDN it is open for the public, but its filename is composed out of its
original filename concatenated with a part of the MD5 hash of the content of the file, which makes
it an extremely hard process to access the file on the CDN without access to the original file or a
reference to the URI.

\subsection{Problem description}
As the current system does not offer proper security measurements, is lacking a lot of features that
is needed and does not scale very well, a new system should be developed. This work is about
examining a way of implementing Copy-on-Write in a high level system like this, which should solve
the scalability problem and make it possible to implement wanted features like snapshots, cloning
and concurrent modifications of content.

\section{Model}
The model for this work should make sure that data can not be accessed by unauthorized users in the
Perius system (See section TODO: fill in). There could also be another model done to make sure that
content can not be accessed by unauthorized viewers once the content is uploaded to the CDN, but as
that should already have been thoroughly checked by the CDN providers this work can focus on the
internal users of the management system.

\subsection{Entities}
%\subsubsection*{snapshot} for not having section numbers
\begin{figure}[htp] \centering{
    \includegraphics[scale=0.8]{relation.pdf}}
    \caption{High Level Entity Relationships}
    \label{fig:relation}
\end{figure}

@@

\subsection{JPF}
Java path finder was used to show that the model and plan of how to build the system was sound. The
model was built in Java with the objective of being as reduced and simple as possible, without
loosing any of the cases that needed to be covered by the model checker. As the users are mainly
going to be handled by external systems they were not included in the model.

@@

\newpage
\begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
elapsed time:       14:26:53
states:             new=160853259,
                    visited=451102505,
                    backtracked=611955764,
                    end=21640
search:             maxDepth=380,
                    constraints=0
choice generators:  thread=160853255 
                    (signal=0,
                    lock=3603938,
                    sharedRef=146989208,
                    threadApi=3,
                    reschedule=10260106), 
                    data=0

@@

\section{Snapshot systems}
\subsection{Copy-on-Write}
\label{sec:copy-on-write}
To efficiently create snapshots of a system Copy-on-Write can be used to make it possible to create
snapshots in O(1)\cite{BTRFS}, this is due to the fact that to create a snapshot in a system using
Copy-on-Write you only need to reference the current nodes in the tree and make sure that they are
not removed, see Figure~\ref{fig:btrfs_tree}.

As the persistent storage, used in this implementation (Section~\ref{persistent_storage}), does not
implement transactions or locks a lot of different problems can occur when several clients are
working on the same data set at the same time. Such problems could be race conditions and
determining the happened-before relation. In this work this problem is solved by implementing
Copy-on-Write.

\subsection{Full Copy}

\subsection{Comparison of Copy-on-Write system implementations}
\subsubsection{BTRFS}
Btrfs is a B-tree file system for Linux which makes use of Copy-on-Write to make it able to do
efficient writeable snapshots and clones. It also supports cloning of subtrees without having to
actually copy the whole subtree, this is due to the Copy-on-Write effect. As several nodes in the
tree can refer to the same node each node keeps track of how many parents it has by a reference
counter so that the node can be deallocated once the node does not have any parents any more. The
reference counter is not stored in the nodes themselves but rather in a separate data structure so
that a nodes counter can be modified without modifying the node itself and therefore eludes the
Copy-on-Write that would have to occur.

\subsubsection{Mach kernel}
In the mid 80s when the development of the Mach kernel started, there was problems with that
physically copying memory was too slow. Too minimise the copying of memory, Copy-on-Write was
implemented. It was implemented so that virtual copy operations could be done and so that tasks
could share read-write memory~\cite{MACH}.

TODO: Insert more systems here
\section{Implementation}
\subsection{Perius}
Perius is the implementation that was done to solve the problem at hand at Uprise. Perius has a
back-end written in Scala and a front-end written in ES6 (Javascript), but they are both
interchangeable.  The back-end has a REST API running, which is how the front-end communicates with
the back-end.

TODO: Picture of the front-end

\subsection{Snapshot functionality}
In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in
Figure~\ref{fig:btrfs_tree}. As Perius does not have the tree structure pre-built and each node is
instead stored in a flat storage space, such operation would be too computationally expensive as
trees would have to be merged when collisions occur, due to the non-blocking nature of the
application. Instead this implementation makes a full copy of the meta-data of the tree, but still
refers to the same binary files until they are changed, which results in the creation of a new node.

\subsection{Resulting system}
This implementation is far from as efficient as the other Copy-on-Write systems described in
Section~\ref{sec:copy-on-write} in most aspects, but more efficient in some. As the implementation
is built upon MongoDB as persistent storage and not a pure tree structure, single nodes can be
fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)),
where n is the number of nodes in the subtree.

\subsubsubsection{API}
REST was chosen as only basic CRUD operations needs to be performed and because the BSON format
which is used in MongoDB is almost identical ~\cite{BSON} to the standardised JSON format which is
used by RESTful services~\cite{JSON}.

  \item projects/\{id\}/containers
      \subitem POST - create new container in a specific project 
  \item projects/\{id\}/containers/\{id\}
      \subitem GET - get specific container in a specific project
      \subitem PUT - update existing container
      \subitem DELETE - delete existing container
  \item projects/\{id\}/containers/\{id\}/content
      \subitem POST - create new content in a specific container 
  \item projects/\{id\}/containers/\{id\}/content/\{id\}
      \subitem GET - get specific content in a specific container
      \subitem PUT - update existing content in a specific container
      \subitem DELETE - delete existing content in a specific container
\end{itemize}

@@

%What to write about
%Battle Binary
%Security features of CDN's
%What is needed by the logic, like what happens after two consecutive deletes
%
%What is needed
%* Security layers
%* snapshots
%* Virtual file structure
%* Versioning of content
%* Multi project support
%* Auth and audit logs
%* Users

>>>>

% useful (wrappers for) math symbols:
\newcommand{\Cardinality}[1]{\left\lvert#1\right\rvert}
\newcommand{\Ceiling}[1]{\left\lceil#1\right\rceil}
\newcommand{\Floor}[1]{\left\lfloor#1\right\rfloor}
\newcommand{\Iff}{\Leftrightarrow}
\newcommand{\Implies}{\Rightarrow}
\newcommand{\Intersect}{\cap}
\newcommand{\Sequence}[1]{\left[#1\right]}
\newcommand{\Set}[1]{\left\{#1\right\}}
\newcommand{\SetComp}[2]{\Set{#1\SuchThat#2}}
\newcommand{\SuchThat}{\mid}
\newcommand{\Tuple}[1]{\langle#1\rangle}
\newcommand{\Union}{\cup}

\newcommand{\fix}{\colorbox{yellow!30}{TODO:}}

\usetikzlibrary{positioning,shapes,shadows,arrows}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords: }} #1}

%}}}
\title{\textbf{Designing a Virtual Security Layer for Cloud Content}}
\author{Lukas Klingsbo}

@@

\pagenumbering{roman}
\setcounter{page}{1}

@@

%}}}
\begin{abstract}
    \fix Abstract

@@

%}}}
\section{Introduction}
Developing larger projects containing static content usually involves using a Content Distribution
Network to be able to scale to a large user base. The commercial Content Distribution Networks are
usually fairly easy to use, the content that is to be used in a project is usually simply uploaded
and then distributed over the globe when the public requests it. For secret content this can be a
problem and an inconvenience, and that is what this thesis is about. This work examines ways of
enforcing virtual access control on content and groups of content, in the form of folders and
snapshots. A system was developed to make the underlying theory work in practice.

The research question that this report answers is how and whether it is practically feasible to use
Copy-on-Write for a high-level system like the one that is implemented.

\fix Clarify problem definition and research questions

\subsection{Prior Work}
\subsubsection{Copy-on-Write}
This work relies heavily on the Copy-on-Write principle, which was founded and used in the Mach
kernel~\cite{COPYONWRITE}, as it can be used to efficiently create snapshots and help solving
concurrency problems that otherwise can occur.

Copy-on-Write is used in for example virtual memory management systems~\cite{VIRTCOW}, for snapshot
and as an optimisation technique for objects and types in several programming
languages\cite{LANGCOW}.

Its principle is that when processes or nodes share data in between each other, the data is not
copied until one of the processes does changes to it. This is an optimisation as the processes does
not have to send or copy all of the related data that is in memory, rather they only have to send
pointers to the data. After many Copy-on-Write's a complex tree structure can be built up, but
optimisations can be done to simplify that structure~\cite{COPYONWRITE2}.

\fix Polish and extend

\newpage
\section{Related Terminology}
\subsection{Abbreviations}
\subsubsection{JPF}
Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
licence, which made more people contribute to the project. JPF is usually used for doing model
checking of concurrent programs to easily find for example race conditions and dead locks.

\subsection{Terms}
\subsubsection{Snapshot}
A snapshot is a way to record the full state of a system at a specific time. The term comes from
photography where a photo can be seen as the state of what the photo is of, at a certain time.
Snapshots should not be confused with full copies of a system as full copies can be used as backups
meanwhile snapshots are not very effective means of backups in the case of data corruption. It is
not effective against data corruption as snapshots usually still refer to unchanged data that is
still a part of the system~\cite{SNAPSHOT}.

\section{Model}
The model for this work should show how the data can not be accessed or modified by unauthorized
users and how the integrity of the data is always kept in the Perius system.

There could also be another relevant model done to show that content can not be accessed by
unauthorized viewers once the content is uploaded to a CDN, but as that should already have been
thoroughly checked by the CDN providers this work can focus solely on the internal users and content
of the management system.

\subsection{Related work}

\subsection{Approach}

\subsection{Design}
\fix Mathematical formal description of how the system works

\subsection{Findings}

\newpage
\section{Implementation}
\subsection{Background}
\subsubsection{About Uprise}
Uprise is a company based in Uppsala, Sweden. It is an EA studio focussing on creating great gaming
experiences, which means that they are not focussed on the actual gameplay which other EA studios
like DICE is.

\subsubsection{The current system}
Today a system called battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files,
mostly images, to content delivery networks. The current system does not make use out of the
security features that the CDN's are offering, instead it uses a form of security by obscurity. When
a file is uploaded to a CDN it is open for the public, but its filename is composed out of its
original filename concatenated with a part of the MD5 hash of the content of the file, which makes
it an extremely hard process to access the file on the CDN without access to the original file or a
reference to the URI.

\subsubsection{Problem description}
As the current system does not offer proper security measurements, is lacking a lot of features that
is needed and does not scale very well, a new system should be developed. This work is about
examining a way of implementing Copy-on-Write in a high level system like this, which should solve
the scalability problem and make it possible to implement wanted features like snapshots, cloning
and concurrent modifications of content.

\subsection{Related Technologies}
\subsubsection{React}
React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and
the browser's, this makes it able to efficiently update dynamic web pages after a change of state
through comparing the old virtual DOM with the resulting virtual DOM after the state change and then
only update the browser's DOM according to the delta between the virtual DOMs~\cite{REACT}. React
can be seen as the system for handling views in front-ends implementing a MVC
(Model-View-Controller) architecture.

@@

@@

\subsection{Snapshot functionality}
\fix Structure to compare snapshot systems and conclude how Perius snapshot system was designed
\subsubsection{Copy-on-Write}
\label{sec:copy-on-write}
To efficiently create snapshots of a system Copy-on-Write can be used to make it possible to create
snapshots in O(1)\cite{BTRFS}, this is due to the fact that to create a snapshot in a system using
Copy-on-Write you only need to reference the current nodes in the tree and make sure that they are
not removed, see Figure~\ref{fig:btrfs_tree}.

As the persistent storage, used in this implementation (Section~\ref{persistent_storage}), does not
implement transactions or locks a lot of different problems can occur when several clients are
working on the same data set at the same time. Such problems could be race conditions and
determining the happened-before relation. In this work this problem is solved by implementing
Copy-on-Write.
\fix Move last paragraph

\subsubsection{Full Copy}
\fix Maybe a short description of full copy, as opposed to copy-on-write

\subsubsection{Comparison of Copy-on-Write system implementations}
\subsubsubsection{BTRFS}
Btrfs is a B-tree file system for Linux which makes use of Copy-on-Write to make it able to do
efficient writeable snapshots and clones. It also supports cloning of subtrees without having to
actually copy the whole subtree, this is due to the Copy-on-Write effect. As several nodes in the
tree can refer to the same node each node keeps track of how many parents it has by a reference
counter so that the node can be deallocated once the node does not have any parents any more. The
reference counter is not stored in the nodes themselves but rather in a separate data structure so
that a nodes counter can be modified without modifying the node itself and therefore eludes the
Copy-on-Write that would have to occur.

\subsubsubsection{Mach kernel}
In the mid 80's when the development of the Mach kernel started, there was problems with that
physically copying memory was too slow. Too minimise the copying of memory, Copy-on-Write was
implemented. It was implemented so that virtual copy operations could be done and so that tasks
could share read-write memory~\cite{MACH}.

\fix Insert more systems here

\subsubsection{Snapshot functionality of Perius}
In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in
Figure~\ref{fig:btrfs_tree}. As Perius does not have the tree structure pre-built and each node is
instead stored in a flat storage space, such operation would be too computationally expensive as
trees would have to be merged when collisions occur, due to the non-blocking nature of the
application. Instead this implementation makes a full copy of the meta-data of the tree, but still
refers to the same binary files until they are changed, which results in the creation of a new node.

\fix Relate section more to comparison

\subsection{Resulting system}
\subsubsection{Perius}
Perius is the implementation that was done to solve the problem at hand at Uprise. Perius has a
back-end written in Scala and a front-end written in ES6 (Javascript), but they are both
interchangeable.  The back-end has a REST API running, which is how the front-end communicates with
the back-end.

\fix Picture of the front-end

\subsubsection{Copy-on-Write}
\fix Rename section and move

This implementation is far from as efficient as the other Copy-on-Write systems described in
Section~\ref{sec:copy-on-write} in most aspects, but more efficient in some. As the implementation
is built upon MongoDB as persistent storage and not a pure tree structure, single nodes can be
fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)),
where n is the number of nodes in the subtree.

\subsubsubsection{API}
In this work a RESTful API was implemented and used for back-end $\Leftrightarrow$ front-end
communication.

REST was chosen as only basic CRUD operations needs to be performed and because the BSON format
which is used in MongoDB is almost identical ~\cite{BSON} to the standardised JSON format which is
used by RESTful services~\cite{JSON}.

  \item projects/\{id\}/containers
      \subitem POST - create new container in a specific project 
  \item projects/\{id\}/containers/\{id\}
      \subitem GET - get specific container in a specific project
      \subitem PUT - update existing container
      \subitem DELETE - delete existing container
\end{itemize}

\section{JPF}
\fix Not sure where to put this section or if it's even relevant any more, maybe skip JPF and put
the entities under implementation -> design
\subsection{Entities}
\begin{figure}[htp] \centering{
    \includegraphics[scale=0.8]{relation.pdf}}
    \caption{High Level Entity Relationships}
    \label{fig:relation}
\end{figure}

@@

\subsection{Execution}
Java path finder was used to show that the model and plan of how to build the system was sound. The
model was built in Java with the objective of being as reduced and simple as possible, without
loosing any of the cases that needed to be covered by the model checker. As the users are mainly
going to be handled by external systems they were not included in the model.

@@

\begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
elapsed time:       14:26:53
states:             new=160853259,
                    visited=451102505,
                    backtracked=611955764,
                    end=21640
search:             maxDepth=380,
                    constraints=0
choice generators:  thread=160853255 
                    (signal=0,
                    lock=3603938,
                    sharedRef=146989208,
                    threadApi=3,
                    reschedule=10260106), 
                    data=0

@@

@@

%What to write about
%Battle Binar
y
%Security features of CDN's
%What is needed by the logic, like what happens after two consecutive deletes
%
%What is needed
%* Security layers
%* snapshots
%* Virtual file structure
%* Versioning of content
%* Multi project support
%* Auth and audit logs
%* Users
